Deep Reinforcement Learning for Conservation Decisions

### 2021-04-28 manuscript/appendix.Rmd
data example

```{r knitr, include = FALSE}
knitr::opts_chunk$set(echo=TRUE, message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 4, cache = TRUE)
ggplot2::theme_set(ggplot2::theme_bw())

@@

@@

noise_std = 0.4805935357322933
action_noise = sb3$common$noise$OrnsteinUhlenbeckActionNoise(mean = np$zeros(1L),
                                            sigma = noise_std * np$ones(1L))

model = sb3$TD3('MlpPolicy', 
            env, 
            verbose = 0, 
            seed = seed,
            "gamma"= 0.995,
            "learning_rate"=  8.315382409902049e-05,
            "batch_size"= 512L,
            "buffer_size"= 10000L,
            "train_freq"= 1000L,
            "gradient_steps"= 1000L,
            "n_episodes_rollout"= -1L,
            "action_noise"= action_noise,
            "policy_kwargs"= list("net_arch"= c(64L,64L)))

>>>>

```{r knitr, include = FALSE}
knitr::opts_chunk$set(echo=TRUE, message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 4, cache = FALSE)
ggplot2::theme_set(ggplot2::theme_bw())

@@

## Applying an RL agent to real data

We use historical data from Argentine Hake to illustrate how an RL agent might be applied in practice.

Historical biomass and catch data for Argentine Hake can be found in the R.A. Myers Legacy Stock Assessment Database [@ramlegacy].
We will use the agent we have just trained using the `gym_fishing` simulator above,

```{r}
# Simulate management under the trained agent
env <- gym$make("fishing-v1", r = 1.0379274,	K = 1.197693, sigma = 0.1121662)
td3 <- sb3$TD3$load("cache/td3")

```{r}
hake <- read_csv("../data/hake.csv")
x0 <- hake %>%
 filter(year == min(year)) %>%
  pull(scaled_biomass)
Tmax <- 15
years <-1986:2000

td3_state <- td3_action <- numeric(Tmax)
td3_state[1] <- x0
td3_action[1] <- NA

Our hindcast considers the scenario of managing catch quotas using TD3, beginning in 1986. 
We compare both the anticipated harvest and resulting stock biomass to the historically recorded harvest and biomass.
This shows that quotas set by TD3 would have been lower than historical values throughout the late 80s and early 90s, which
saw a sharp decline in estimated biomass of the Argentine Hake.
For the same or higher estimated stock sizes from the historical record, we see that TD3 would recommend a lower quota.
We also see that under TD3 the stock recovers sufficiently to predict higher quotas by the second half of the 90s than were
being set in the now-depleted stock.  
Whether the TD3 quotas would have truly been sufficient to permit recovery in stock sizes shown in the simulation clearly depends on the accuracy of the underlying simulation.
However, the fact that TD3 would set a lower quota than the historical management given the same or even higher estimate of stock size is independent of the Hake simulation.
Combined with the observed historical declines, this provides some evidence that the RL agent could have decreased or avoided the over-harvesting in this case.

```{r}
env$reset()
## represent the initial state size in the 'rescaled' state space.
state <- env$get_state( x0 )

for(i in 2:Tmax){

  # RL-recommended harvest action:
  out <- td3$predict(state)
  action <- out[[1]]

  # Record state and resulting action
  td3_state[i] <- env$get_fish_population(state)
  td3_action[i] <- env$get_quota(action)

  # Implement RL-recommended harvest. 
  result <- env$step(action)
  state <- result[[1]]

  #state <- env$get_state(hake$scaled_biomass[i+1])
}

df <- bind_rows(
  tibble(year = years, state = td3_state, action = td3_action, model = "TD3"),
  tibble(year = years, state = hake$scaled_biomass, action = hake$scaled_catch, model = "historical"))

df %>% ggplot(aes(year, state, col=model)) + geom_line() + geom_point(aes(year, action, col=model))

@@

```{r}
noise_std = 0.4805935357322933

OU <- sb3$common$noise$OrnsteinUhlenbeckActionNoise
action_noise = OU(mean = np$zeros(1L),  sigma = noise_std * np$ones(1L))

model = sb3$TD3('MlpPolicy', 
            env, 
            verbose = 0, 
            seed = 42L,
            "gamma"= 0.995,
            "learning_rate"=  8.315382409902049e-05,
            "batch_size"= 512L,
            "buffer_size"= 10000L,
            "train_freq"= 1000L,
            "gradient_steps"= 1000L,
            "n_episodes_rollout"= -1L,
            "action_noise"= action_noise,
            "policy_kwargs"= list("net_arch"= c(64L,64L)))

### 2021-04-28 manuscript/manuscript.Rmd
data example

In contrast to other ML methods, RL does not require massive data sets for training.  
Despite the expansion in data collection from remote sensors, citizen science, and other efforts, ecological data remains very sparse on many necessary details.
Moreover, no amount of data can train supervised or unsupervised ML approaches can predict a future which has no analog in the historical record.
Mechanistic, process-based models remain our most powerful tool to explore scenarios of dynamic ecosystems in changing climates.
Methods from _optimal control theory_ have long been used in conservation and resource economics literature [@Clark1990] to determine optimal policies given such models [@Polasky2011].
However, control theory approaches are also tightly constrained by the "curse of dimensionality" [@Mangel1980; @Marescot2013], limiting their application to only simple, stylized models.
Deep reinforcement learning offers a way forward. 
Deep RL agents are trained not with mountains of data, but by interacting with simulators.
Most recent advances in Deep RL have illustrated algorithms trained by playing arcade game simulators such as Atari BreakOut, or physics-based simulations of robotics movement [@TD3; @SAC; @PPO; @A2C].
Simulating from complex models is far simpler than solving them.
The richest ecosystem models, such as the global circulation models (GCMs) used to generate future climate scenarios [@something], or individual-based models used to predict forest growth [@something] or wildfire spread [@something], are massive simulations. 
Given sufficient computational power, RL can be trained and evaluated across a broad range of possible realistic scenarios well beyond the reach of classical methods.
In this manner, RL would compliment, rather than replace, existing empirical and theoretical research which would continue to form the basis for those simulations.

>>>>

In contrast to other ML methods, RL does not require massive data sets for training, relying instead on model simulations based on the best available ecological data and knowledge of underlying processes.
Despite the expansion in data collection from remote sensors, citizen science, and other efforts, ecological data remains sparse and unevenly sampled [CITE].
Moreover, no amount of data can train supervised or unsupervised ML approaches can predict a future which has no analog in the historical record.
Mechanistic, process-based models remain our most powerful tool to explore scenarios of dynamic ecosystems in changing climates.
Methods from _optimal control theory_ have long been used in conservation and resource economics literature [@Clark1990] to determine optimal policies given such models [@Polasky2011].
However, control theory approaches are also tightly constrained by the "curse of dimensionality" [@Mangel1980; @Marescot2013], limiting their application to only simple, stylized models.
Deep reinforcement learning offers a way forward. 
Deep RL agents are trained not with mountains of data, but by interacting with simulators.
Most recent advances in Deep RL have illustrated algorithms trained by playing arcade game simulators such as Atari BreakOut, or physics-based simulations of robotics movement [@TD3; @SAC; @PPO; @A2C].
Simulating from complex models is far simpler than solving them.
The richest ecosystem models, such as the global circulation models (GCMs) used to generate future climate scenarios [@something], or individual-based models used to predict forest growth [@something] or wildfire spread [@something], are massive simulations. 
Given sufficient computational power, RL can be trained and evaluated across a broad range of possible realistic scenarios well beyond the reach of classical methods.
In this manner, RL would compliment, rather than replace, existing empirical and theoretical research which would continue to form the basis for those simulations.
